Methods to prioritise pop-up active transport infrastructure

### 2020-05-09 article.Rmd
Update article

```{r parameters}
min_pctgov = 0
min_grouped_pctgov = 100
min_grouped_length = 500
city_centre_buffer_radius = 8000
city_centre = tmaptools::geocode_OSM("leeds", as.sf = TRUE)
city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)
h_leeds = hsf[city_centre_buffer, ]
```

```{r preprocess}
# remove motorways
r_original$highway2 = r_original$highway
r_original$highway2 = gsub(pattern = "_link", replacement = "", r_original$highway)
highway_table = table(r_original$highway)
highway_rare = highway_table[highway_table < nrow(r_original) / 100]
highway_remove = names(highway_rare)[!grepl(pattern = "motor|living|ped", x = names(highway_rare))]
r_cleaned = r_original %>% 
  filter(!grepl(pattern = "motorway", x = highway)) %>% 
  mutate(highway2 = case_when(
  highway2 %in% highway_remove ~ "other",
  grepl(pattern = "path|track|trunk", highway2) ~ "other",
  grepl(pattern = "ped|liv", highway2) ~ "pedestrian/living_street",
  TRUE ~ highway2
  )) %>% 
  mutate(maxspeed = case_when(
    maxspeed <= 20 ~ "20 mph or less",
    maxspeed > 20 & maxspeed <= 30 ~ "30 mph",
    maxspeed > 30 ~ "40+ mph",
  )) %>% 
  mutate(pctgov = as.numeric(pctgov)) 
table(r_cleaned$maxspeed)
rj = inner_join(r_cleaned, rtid) %>% 
  mutate(lanes_f = lanespsvforward + lanesforward) %>% 
  mutate(lanes_b = lanespsvbackward + lanesbackward) %>% 
  mutate(`N. lanes` = lanes_f + lanes_b)
cy = r_cleaned %>% filter(highway == "cycleway")
r_central = rj[city_centre_buffer, ]
r_main = r_central %>% 
  filter(grepl(pattern = "cycleway|primary|second|tert", highway2))
r_main_region = rj %>% 
  filter(grepl(pattern = "cycleway|primary|second|tert", highway2))
```

@@

```{r t1, results='asis'}
# rj %>%
#   st_drop_geometry() %>%
#   select(name, highway2, maxspeed, pctgov, width) %>%
#   table1::table1(~ highway2 + pctgov + width | maxspeed, data = ., )
knitr::include_graphics("https://user-images.githubusercontent.com/1825120/81476361-8c094480-9209-11ea-97d4-f4ad09e96d69.png")
```

```{r hospitals, fig.cap="Overview map of input data, showing the main highway types and location of hospitals in Leeds"}
m1 = r_main_region %>%
  mutate(`Highway type` = highway2) %>% 
  mutate(`Cycling potential` = case_when(pctgov < 100 ~ 100, TRUE ~ pctgov)) %>% 
  tm_shape(bbox = st_bbox(city_centre_buffer)) +
  tm_lines(col = "Highway type", palette = c("green", "black", "blue", "grey"),
           lwd = "Cycling potential", scale = 5, lwd.legend = c(100, 200, 500, 1000),
           lwd.legend.labels = c("0 - 100", "100 - 200", "200 - 500", "500+")) +
  tm_shape(h_leeds) + tm_dots(size = 0.5, col = "ParentName", palette = "Dark2", title = "Hospital group") +
  tm_layout(legend.outside = TRUE)
# + tm_text("OrganisationName")
m1
```

@@

```{r gsub, fig.cap="Illustration of geographic subsetting based on distance to the city centre rather than based on location within somewhat arbitrarily shaped city boundaries"}
# nrow(r_central)
# nrow(rj)
square = stplanr::geo_bb(shp = city_centre_buffer)
m1 = tm_shape(r_main_region) +
  tm_lines("maxspeed", palette = "-Set1", lwd = "N. lanes", scale = 5,
           lwd.legend = c(1, 2, 3, 4, 5, 6)) +
  tm_shape(region) + tm_borders() +
  tm_shape(square) + tm_borders(lty = 3) +
  tm_scale_bar() 
# m1
m2 = tm_shape(r_main) +
  tm_lines("maxspeed", palette = "-Set1", lwd = "N. lanes", scale = 5,
           lwd.legend = c(1, 2, 3, 4, 5, 6)) +
  tm_shape(city_centre_buffer) + tm_borders() +
  tm_scale_bar() +
  tm_layout(legend.show = FALSE)
tmap_arrange(m1, m2, nrow = 1)
```

@@

```{r levels}
r_pct_lanes = rj %>% 
  filter(pctgov > min_pctgov) %>% 
  filter((lanespsvforward + lanesforward) > 1 |
      (lanespsvbackward + lanesbackward) > 1
  )
touching_list = st_touches(r_pct_lanes)
g = igraph::graph.adjlist(touching_list)
components = igraph::components(g)
r_pct_lanes$group = components$membership
r_pct_lanes = r_pct_lanes %>% 
  group_by(group) %>% 
  mutate(group_length = sum(length)) %>% 
  mutate(pctgov_mean = mean(pctgov, na.rm = TRUE)) %>% 
  filter(pctgov_mean > min_grouped_pctgov)
r_pct_lanes = r_pct_lanes %>% filter(group_length > min_grouped_length)
r_filter_before_grouping = rj %>% 
  filter(pctgov > min_pctgov) %>% 
  filter((lanespsvforward + lanesforward) > 1 |
           (lanespsvbackward + lanesbackward) > 1
  ) %>% 
  filter(pctgov > min_grouped_pctgov) %>% 
  filter(length > 100)
tmap_mode("plot")
m1 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
  tm_shape(r_filter_before_grouping) + tm_lines() +
  tm_layout(title = "Filter then group (length > 100, cycling_potential > 100)")
m2 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
  tm_shape(r_pct_lanes) + tm_lines() +
  tm_layout(title = "Group then filter (length > 500, cycling_potential > 100)")
ma = tmap_arrange(m1, m2)
ma
```

>>>>

```{r parameters}
min_cycling_potential = 0
min_grouped_cycling_potential = 100
min_grouped_length = 500
city_centre_buffer_radius = 8000
city_centre = tmaptools::geocode_OSM("leeds", as.sf = TRUE)
city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)
h_leeds = hsf[city_centre_buffer, ]
```

```{r preprocess}
# remove motorways
r_original$highway_type = r_original$highway
r_original$highway_type = gsub(pattern = "_link", replacement = "", r_original$highway)
highway_table = table(r_original$highway)
highway_rare = highway_table[highway_table < nrow(r_original) / 100]
highway_remove = names(highway_rare)[!grepl(pattern = "motor|living|ped", x = names(highway_rare))]
r_cleaned = r_original %>% 
  filter(!grepl(pattern = "motorway", x = highway)) %>% 
  mutate(highway_type = case_when(
  highway_type %in% highway_remove ~ "other",
  grepl(pattern = "path|track|trunk", highway_type) ~ "other",
  grepl(pattern = "ped|liv", highway_type) ~ "pedestrian/living_street",
  TRUE ~ highway_type
  )) %>% 
  mutate(maxspeed = case_when(
    maxspeed <= 20 ~ "20 mph or less",
    maxspeed > 20 & maxspeed <= 30 ~ "30 mph",
    maxspeed > 30 ~ "40+ mph",
  )) %>% 
  mutate(cycling_potential = as.numeric(pctgov)) 
# table(r_cleaned$maxspeed)
rj = inner_join(r_cleaned, rtid) %>% 
  mutate(lanes_f = abs(lanespsvforward) + abs(lanesforward)) %>% 
  mutate(lanes_b = abs(lanespsvbackward) + abs(lanesbackward)) %>% 
  mutate(n_lanes_numeric = lanes_f + lanes_b) %>% 
  mutate(n_lanes = as.character(n_lanes_numeric))
rj$n_lanes_numeric[rj$n_lanes_numeric <= 0] = 1 
rj$n_lanes[rj$n_lanes == "0"] = "1"
rj$n_lanes[rj$n_lanes == "4"] = "4+"
rj$n_lanes[rj$n_lanes == "5"] = "4+"
rj$n_lanes[rj$n_lanes == "6"] = "4+"
cy = r_cleaned %>% filter(highway == "cycleway")
r_central = rj[city_centre_buffer, ]
r_main = r_central %>% 
  filter(grepl(pattern = "cycleway|primary|second|tert", highway_type))
r_main_region = rj %>% 
  filter(grepl(pattern = "cycleway|primary|second|tert", highway_type))
```

@@

```{r t1, results='asis'}
# rj %>%
#   st_drop_geometry() %>%
#   # select(name, highway_type, maxspeed, cycling_potential, width) %>%
#   table1::table1(~ highway_type + cycling_potential + width + n_lanes | maxspeed, data = ., )
knitr::include_graphics("https://user-images.githubusercontent.com/1825120/81476961-c32d2500-920c-11ea-8430-94c3afc2e27d.png")
```

```{r hospitals, fig.cap="Overview map of input data, showing the main highway types and location of hospitals in Leeds"}
m1 = r_main_region %>%
  mutate(`Highway type` = highway_type) %>% 
  mutate(`Cycling potential` = case_when(cycling_potential < 100 ~ 100, TRUE ~ cycling_potential)) %>% 
  tm_shape(bbox = st_bbox(city_centre_buffer)) +
  tm_lines(col = "Highway type", palette = c("green", "black", "blue", "grey"),
           lwd = "Cycling potential", scale = 5, lwd.legend = c(100, 200, 500, 1000),
           lwd.legend.labels = c("0 - 100", "100 - 200", "200 - 500", "500+")) +
  tm_shape(h_leeds) + tm_dots(size = 0.5, col = "ParentName", palette = "Dark2", title = "Hospital group") +
  tm_layout(legend.outside = TRUE)
# + tm_text("OrganisationName")
m1
```

@@

```{r gsub, fig.cap="Illustration of geographic subsetting based on distance to the city centre rather than based on location within somewhat arbitrarily shaped city boundaries"}
# nrow(r_central)
# nrow(rj)
square = stplanr::geo_bb(shp = city_centre_buffer)
m1 = tm_shape(r_main_region) +
  tm_lines("maxspeed", palette = "-Set1", lwd = "n_lanes_numeric", scale = 5,
           lwd.legend = c(1, 2, 3, 4, 5, 6), title.lwd = "Number of lanes") +
  tm_shape(region) + tm_borders() +
  tm_shape(square) + tm_borders(lty = 3) +
  tm_scale_bar() 
# m1
m2 = tm_shape(r_main) +
  tm_lines("maxspeed", palette = "-Set1", lwd = "n_lanes_numeric", scale = 5,
           lwd.legend = c(1, 2, 3, 4, 5, 6), title.lwd = "Number of lanes") +
  tm_shape(city_centre_buffer) + tm_borders() +
  tm_scale_bar() +
  tm_layout(legend.show = FALSE)
tmap_arrange(m1, m2, nrow = 1)
```

@@

```{r levels}
r_pct_lanes = rj %>% 
  filter(cycling_potential > min_cycling_potential) %>% 
  filter((lanespsvforward + lanesforward) > 1 |
      (lanespsvbackward + lanesbackward) > 1
  )
touching_list = st_touches(r_pct_lanes)
g = igraph::graph.adjlist(touching_list)
components = igraph::components(g)
r_pct_lanes$group = components$membership
r_pct_lanes = r_pct_lanes %>% 
  group_by(group) %>% 
  mutate(group_length = sum(length)) %>% 
  mutate(cycling_potential_mean = mean(cycling_potential, na.rm = TRUE)) %>% 
  filter(cycling_potential_mean > min_grouped_cycling_potential)
r_pct_lanes = r_pct_lanes %>% filter(group_length > min_grouped_length)
r_filter_before_grouping = rj %>% 
  filter(cycling_potential > min_cycling_potential) %>% 
  filter((lanespsvforward + lanesforward) > 1 |
           (lanespsvbackward + lanesbackward) > 1
  ) %>% 
  filter(cycling_potential > min_grouped_cycling_potential) %>% 
  filter(length > 100)
tmap_mode("plot")
m1 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
  tm_shape(r_filter_before_grouping) + tm_lines() +
  tm_layout(title = "Filter then group (length > 100, cycling_potential > 100)")
m2 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
  tm_shape(r_pct_lanes) + tm_lines() +
  tm_layout(title = "Group then filter (length > 500, cycling_potential > 100)")
ma = tmap_arrange(m1, m2)
ma
```
